\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{verbatim}
\usepackage{footnote}
\usepackage{xr}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}


\newtheorem{constraint}{\textbf{Constraint}}

\newenvironment{defn}
{\vspace{12pt} \noindent {\textit{meaning:}}} \\

\begin{document}

\title{Paper Title*\\
{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
should not be used}
\thanks{Identify applicable funding agency here. If none, delete this.}
}

\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
}

\maketitle

\begin{abstract}
Floorplanning is a mandatory step in the design of hardware accelerators on FPGA platforms consisting in placing the hardware modules to specific FPGA areas. In particular, if the design makes use of partial reconfiguration (PR), the overall system performance is greatly influenced by the allocation strategy. Existing tools for PR, which are provided by FPGA vendors, allow floorplanning to be performed manually. This method besides being highly time-consuming, demands a good knowledge of the architecture and organization of the hardware for an optimal floorplan. To solve this problem, this paper presents an automated floorplanner based on Mixed Integer Programming (MILP). The most common approach by the state of the art automated MILP (or other optimization method) based floorplanners is to apply MILP (or other optimization methods) for finding the optimal placement from a pre-enumerated list of possible placements. Meanwhile, in this work, the floorplanning problem was directly solved as an optimization problem by modeling the FPGA resources as a set of MILP constraints along with the standard PR related constraints. The performance of the proposed approach was tested by using a synthethic benchmark suite. Its performance was also compared with the state of the art MILP based floorplanners for the average execution time. Finally, the floorplanner was used in a real case study. This approach was observed to improve the execution time reported in the state of the art by orders of magnitude.
\end{abstract}

\begin{IEEEkeywords}
component, formatting, style, styling, insert
\end{IEEEkeywords}

\input{introduction}
\input{related_works}
\input{model}
\input{problem}
\input{milp_formulation}
\input{exps}

\end{document}
