\section{Introduction}
\textbf{par 1} \\
\textit{what is floorplanning and how is it applied (what's its role) in the context of PR, what is the benefit of having a good floorplan or a floorplan at all} \\

\textbf{par 2} \\
\textit{How is fp done currently and what are the shortcomings of those methods} \\

\textbf{par 3} \\
\textit{In short what did I do and what are my contributions (clearly stated)} \\

\textbf{par 4} \\
\textit{outline of the work} \\

Floorplanning is one of the major challenges in the field of dynamic parital reconfiguration. The placement of the static and reconfigurable slots on the FPGA fabric must satisfy the application requirements set by the application designer while also respecting the technological constraints set by the manufacturer. The conventional approach for an automated generation of FPGA floorplans usually involves two steps. First for each slot, all the possible rectangular slots that satisfy the resources requirement of the slot are enumerated. This is done by starting a scan on the fpga fabric from the bottom left corner and lisitng all the rectangles that contain all the necessary resources for the respective slots. Then some sort of heuristics/optimization is applied to choose the optimal ones from the set of possible slots. This approach has many problems \{\textit{to be listed later}\}\\\\

Our approach instead focuses on applying the optimization process on a lower level of abstraction of the fpga fabric i.e., rather than applying optimization to select the most optimal one from a set of pre-scanned slots, we modeled the different types of resources on the fpga and their distribution along with forbidden regions as a set of constraints and added these constraints to the predefined constraints related to dpr. \\

Let us consider a floorplanning example where we have to make a floorplan for two slots S$_1$ and S$_2$ on the FPGA fabric. Each slot has resource requirements denoted as \{D$_1$, B$_1$, C$_1$\} and \{D$_2$, B$_2$, C$_2$\} where D, B and C represent DSP, BRAM and CLB respectively. \\ Our proposed system takes as an input in the resource requirement of each slot and a description of the resource distribution of the FPGA fabric and it returns the placement coordinates of the slots on the fpga fabric. \\
A slot is represented using 4 parameters i.e. the two bottom left coordinates and the width and the height of the slot. In our considered example the slots S$_1$ and S$_2$ are represented as (x$_1$, y$_1$, w$_1$, h$_1$) and (x$_2$, y$_2$, w$_2$, h$_2$). A forbidden region is also represented as a slot hence a forbidden region F$_i$ can be represented as {fx$_i$, fy$_k$, fw$_i$, fy$_i$}  \\ 


